---
title: "R-tistic"
subtitle: "Data Visualisation with ggplot2"
date: March 16, 2017
author: Lars Schoebitz </br>r-tistic@lse.de
output:
  xaringan::moon_reader:
    css: "r-tistic_theme.css"
    lib_dir: libs
    self_contained: FALSE
    chakra: "remark-lib/remark-latest.min.js"
    nature:
      highlightLanguage: R
      highlightLines: true
    navigation: 
      click: true
---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)
library(tidyverse)
library(knitr)
library(grDevices)
library(Cairo)
library(gapminder)

```

```{r download-remark-bib, include=FALSE}

# xaringan::summon_remark(version = "latest", to = "remark-lib/")

```

```{r make-pdf, message=FALSE, warning=FALSE, include=FALSE}

## print slides as PDF document

# system("~/decktape-1.0.0/phantomjs ~/decktape-1.0.0/decktape.js ~/R/R-tistic/aCourses/02-data-visualisation.html ~/R/R-tistic/aCourses/02-data-visualisation.pdf")


```

layout: true
class: theme

---
## ggplot2
***
### Key components

Every ggplot2 plot has three key components:

1. **data**
2. A set of **aesthetic mappings** between variables in the data and visual properties, and
3. At lest one layer which describes how to render each observation. Layers are usually created with a **geom** function.

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot2
***
### Example

- type the code below into the console and hit `↵`

```{r, eval=FALSE}

ggplot(mpg, aes(x = displ, y = hwy)) +  
  geom_point()

```

--

```{r intro, echo=FALSE, fig.height=3, fig.width=8, dev='svg', cache=TRUE}

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() 

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot2
***
### What does the function do?

--

1. Data: `mpg` dataset, which is build into ggplot2.
2. Aesthetic mapping: engine size mapped to x position, fuel economy to y position.
3. Layer: points.

--

Also being referred to as a scatterplot.

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Exercises

- type the code below into the console and hit `↵`

```{r, eval=FALSE}

ggplot(gapminder, aes(country, lifeExp)) +  
  geom_point()

```

--

- what does the plot show?

```{r intro-2, echo=FALSE, fig.height=3, fig.width=8, dev='svg', cache=TRUE}

ggplot(gapminder, aes(country, lifeExp)) +  
  geom_point()


```


---
## ggplot 2
***
### Exercises


```{r, eval=FALSE}

ggplot(gapminder, aes(country, lifeExp)) +  
  geom_point()

```


```{r ggplot-1, echo=FALSE, fig.height=5, fig.width=8, dev='svg', cache=TRUE}

ggplot(gapminder, aes(x = lifeExp, y = country)) +  
  geom_point()

```

---
## ggplot 2
***
### Exercises

```{r, eval=FALSE}
ggplot(gapminder, aes(x = country, y = lifeExp)) +  
  geom_point() +
  coord_flip()
```


```{r intro-3, echo=FALSE, fig.height=4.5, fig.width=8, dev='svg', cache=TRUE}

ggplot(gapminder, aes(x = country, y = lifeExp)) +  
  geom_point() +
  coord_flip()

```

---
## ggplot 2
***
### Exercises: Make a scatterplot

1. Data: `diamonds` dataset, which is build into ggplot2.
2. Aesthetic mapping: carat mapped to x position, price mapped to y position.
3. Layer: points.

--

```{r, eval=FALSE}

ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point()

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Exercises: Make a scatterplot

```{r intro-4, echo=FALSE, fig.height=4, fig.width=8, dev='CairoSVG', cache = TRUE}

ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point()

```

---
## ggplot 2
***
### Aesthetic Attributes

- add colour, shape or size to `aes()` 

```{r, eval=FALSE}

ggplot(mpg, aes(x = displ, y = cty, colour = class)) +
  geom_point()

```

--

```{r mpg-aes-colour, echo=FALSE, fig.height=3, fig.width=8, dev='CairoSVG'}

ggplot(mpg, aes(x = displ, y = cty, colour = class)) +
  geom_point() 

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Aesthetic Attributes

- map shape instead of colour
- then map size instead of colour
- what happens?

```{r, eval=FALSE}

ggplot(mpg, aes(x = displ, y = cty, shape = class)) +
  geom_point() 

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Warnings

```{r mpg-aes-shape, fig.height=2, fig.width=8, warning=TRUE, dev='CairoSVG'}

ggplot(mpg, aes(x = displ, y = cty, shape = class)) +
  geom_point() 

```

---
## ggplot 2
***
### Warnings

```{r mpg-aes-size, fig.height=3, fig.width=8, warning=TRUE, dev='CairoSVG', cache=TRUE}

ggplot(mpg, aes(x = displ, y = cty, size = class)) +
  geom_point() 

```

---
# Variables
***
### Discrete, continous and categorical

- colour and shape work well with categorical variables
- size works well for continous variables
- ...

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Exercise

- type the code below into the console and hit `↵`
- compare the two plots

```{r, eval=FALSE}

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = blue))

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(colour = blue)

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Exercise

```{r mpg-aes-colour-geom, echo=FALSE, fig.height=4, fig.width=9, warning=TRUE, cache=TRUE, dev='CairoSVG'}

library(gridExtra)

g1 <- ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = "blue"))

g2 <- ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(colour = "blue")

grid.arrange(g1, g2, ncol = 2)


```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Facetting

- another technique to display additional categorical variables on a plot
- type the code below into the console and hit `↵`

```{r, eval=FALSE}

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  facet_wrap(~class)

```

- what can you see?

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Facetting

- what can you see?

```{r mpg-facetting, echo=FALSE, fig.height=4, fig.width=9, warning=TRUE, cache=TRUE, dev='CairoSVG'}

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  facet_wrap(~class)

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Plot Geoms

- geom_smooth
- geom_boxplot
- geom_histogram
- geom_bar
- geom_path
- geom_line

---
## ggplot 2
***
### geom_smooth()

Make a scatterplot using the `mpg` dataset with the aesthetic mappings engine size and fuel economy position. Then add a smoother.

--

```{r, eval=FALSE, message=FALSE, warning=FALSE}

ggplot(mpg, aes(x = displ, y = hwy)) +  
  geom_point() +
  geom_smooth()

```

--

```{r mpg-smooth, echo=FALSE, fig.height=3, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE, dev='svg'}

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### geom_smooth() span

```{r mpg-smooth-span-1, echo=TRUE, fig.height=3, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE, dev='svg'}

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(span = 0.2)

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### geom_smooth() span

```{r mpg-smooth-span-2, echo=TRUE, fig.height=3, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE, dev='svg'}

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(span = 1)

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### geom_smooth() linear model

```{r mpg-smooth-lm, echo=TRUE, fig.height=3, fig.width=8, message=FALSE, warning=FALSE, cache=TRUE, dev='svg'}

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(method = "lm")

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Boxplots and Jittered points

```{r mpg-point-1, echo=FALSE, fig.height=3, fig.width=8, cache=TRUE, dev='svg'}

ggplot(mpg, aes(x = drv, y = hwy)) +
  geom_point() +
  theme_bw(base_size = 24)

```

--

- few unique values (lots of overplotting)
    - geom_jitter
    - geom_boxplot
    - geom_violin
    
.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]
    
---
## ggplot 2
***
### Boxplots, Jittered Points and Violin Plots

```{r mpg-jitter-box-violin, echo=FALSE, fig.height=3, fig.width=8, cache=TRUE, dev='svg'}

g_jitter <- ggplot(mpg, aes(x = drv, y = hwy)) +
  geom_jitter() +
  theme_bw(base_size = 24)

g_boxplot <- ggplot(mpg, aes(x = drv, y = hwy)) +
  geom_boxplot() +
  theme_bw(base_size = 24)

g_viloin <- ggplot(mpg, aes(x = drv, y = hwy)) +
  geom_violin() +
  theme_bw(base_size = 24)

grid.arrange(g_jitter, g_boxplot, g_viloin, nrow = 1)

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Histograms 

```{r, eval=FALSE}

ggplot(mpg, aes(hwy)) +
  geom_histogram() 

```

--

```{r mpg-histogram-1, echo=FALSE, fig.height=3, fig.width=8, cache=TRUE, dev='svg'}

ggplot(mpg, aes(hwy)) +
  geom_histogram() 

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Histograms 

```{r, eval=FALSE}

ggplot(mpg, aes(hwy)) +
  geom_histogram(binwidth = 2.5) 

```

--

```{r mpg-histogram, echo=FALSE, fig.height=3, fig.width=8, cache=TRUE, dev='svg'}

ggplot(mpg, aes(hwy)) +
  geom_histogram(binwidth = 2.5) 

```


.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Frequency Polygons 

```{r, eval=FALSE}

ggplot(mpg, aes(hwy)) +
  geom_freqpoly() 

```

--

```{r mpg-freq-poly, echo=FALSE, fig.height=3, fig.width=8, cache=TRUE, dev='svg'}

ggplot(mpg, aes(hwy)) +
  geom_freqpoly() 

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## ggplot 2
***
### Frequency Polygons and Histograms

```{r, echo=FALSE, fig.height=6, fig.width=12, message=FALSE, warning=FALSE}

library(gridExtra)

g_freq <- ggplot(mpg, aes(displ, colour = drv)) +
  geom_freqpoly(binwidth = 0.5) +
  theme_bw(base_size = 20)

g_hist <- ggplot(mpg, aes(displ, fill = drv)) +
  geom_histogram(binwidth = 0.5)  +
  facet_wrap(~drv, ncol = 1) +
  theme_bw(base_size = 20)

grid.arrange(g_freq, g_hist, ncol = 2)

```

.footnote[
This slide is inspired by content and code directly copied from the book ["ggplot2 - Elegant Graphics for Data Analysis"](http://bit.ly/2muDLCd) by [Hadley Wickhmam](http://hadley.nz/). Use R! Springer International Publishing, 2016. doi:10.1007/978-3-319-24277-4_2.
]

---
## Library
***
### GridExtra

- install package gridExtra
- load package gridExtra into your script

--

```{r, eval=FALSE}

library(gridExtra)

g_freq <- ggplot(mpg, aes(displ, colour = drv)) +
  geom_freqpoly(binwidth = 0.5) +
  theme_bw(base_size = 20)

g_hist <- ggplot(mpg, aes(displ, fill = drv)) +
  geom_histogram(binwidth = 0.5)  +
  facet_wrap(~drv, ncol = 1) +
  theme_bw(base_size = 20)

grid.arrange(g_freq, g_hist, ncol = 2)

```

---
## Next steps
***
### Learn more and have fun!

.big[
[Try R Codeschool](http://bit.ly/2nrWfHQ)
]

.big[
["R for Data Science"](http://bit.ly/wickham_r4ds) by [Hadley Wickhmam](http://hadley.nz/)
]


---
## Final words
***

### Contact
.button1[[email](mailto:r-tistic@lse.de)]
.button1[[website](http://www.lse.de/r-tistic)]
.button1[[twitter](https://twitter.com/Larnsce)]

### Slides
Source code available on [GitHub](http://bit.ly/rtistic_repo)

### References 
RStudio Team (2016). RStudio: Integrated Development for R. RStudio, Inc., Boston, MA URL http://www.rstudio.com/  
Slides created via the R package [xaringan](https://github.com/yihui/xaringan)  
The chakra comes from [remark.js](https://remarkjs.com), [knitr](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com)  




